---
title: 'Project Code'
output:
  html_document:
  df_print: paged
html_notebook: default
pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE)
```

```{r}
# Load required libraries
library(tidyverse)
library(readr)
library(readxl)
library(stringr)
# Set scientific notation preference
options(scipen = 10)
```

```{r}
# Define file paths
file_paths <- list(
  social_capital_county = "social_capital_county.csv",
  world_happiness_report = "WorldHappinessReport2022-Score.csv",
  stats_by_state = "ComprehensiveStatsByState-PopPercent-20000101_20220815.csv",
  world_regions = "WorldRegions.csv",
  climate_change_survey = "climate_change_opinion_survey_2022_aggregated.xlsx",
  primary_energy = "INT-Export-PrimaryEnergy-09-09-2022_23-12-37.csv",
  natural_gas = "INT-Export-NaturalGas-09-09-2022_23-11-09.csv",
  covid = "COVID-2022-11-01.csv",
  emissions = "INT-Export-Emissions-09-09-2022_23-18-25.csv"
)

# Load datasets
social_capital_county <- read_csv(file_paths$social_capital_county)
world_happiness_report <- read_csv(file_paths$world_happiness_report)
stats_by_state <- read_csv(file_paths$stats_by_state)
world_regions <- read_csv(file_paths$world_regions)
climate_change_survey <- read_excel(file_paths$climate_change_survey)
primary_energy <- read_csv(file_paths$primary_energy, skip = 1)
natural_gas <- read_csv(file_paths$natural_gas, skip = 1)
covid <- read_csv(file_paths$covid)
emissions <- read_csv(file_paths$emissions, skip = 1)

#display loaded datasets (uncomment to inspect)
# head(social_capital_county)
# head(world_happiness_report)
# head(world_regions)
# head(natural_gas)
# head(emissions)
```

```{r}
# Splitting the combined column into separate variables
world_regions <- world_regions %>%
  separate(`Country\tRegion\tGlobal South`, into = c("Country", "Region", "Global"), sep = "\t")

# Selecting relevant columns
world_regions <- world_regions %>% select(Country, Region, Global)

# View cleaned dataset
head(world_regions)

# Extracting region-level data (excluding specific categories)
regions <- natural_gas %>%
  filter(is.na(API)) %>%
  filter(!...2 %in% c("Production", "Consumption", "Imports", "Exports", "Reserves")) %>%
  rename(Country = ...2)

# Extracting country-level natural gas data
gas <- natural_gas %>%
  filter(!is.na(API)) %>%
  rename(`Natural Gas Type` = ...2)

# View cleaned datasets
head(regions)
head(gas)

# Replace missing values (denoted as "--") with NA
emissions[emissions == "--"] <- NA

# Extracting region-level emissions data
regions_emissions <- emissions %>%
  filter(is.na(API)) %>%
  filter(...2 != "Emissions") %>%
  rename(Country = ...2)

# Extracting CO2 emissions and converting to numeric format
co2_emissions <- emissions %>%
  filter(...2 == "CO2 emissions (MMtonnes CO2)") %>%
  rename(`Emission Type` = ...2) %>%
  mutate(Region = regions_emissions$Country) %>%
  select(Region, -API) %>%
  mutate_if(is.character, as.double) %>%
  round(3)

# View cleaned dataset
head(co2_emissions)
```

```{r}
#Visualizations

ggplot(co2_emissions, aes(x = Region, y = `2018`)) +
  geom_line() +
  labs(title = "CO2 Emissions by Region (2018)",
       x = "Region",
       y = "CO2 Emissions (MMtonnes)") +
  theme_minimal()
```

```{r}
Tidying Social Capital report:
```{r}
social_capital <- social_capital_county %>% 
  rename(c("County Code" = "county", "County Name" = "county_name"), "LowIncome" = "num_below_p50", "Population" = "pop2018", "EconomicConnectedness" = "ec_county", "Socioeconomic" = "ec_se_county", "High Income"= "ec_grp_mem_county", "Groups EC"="ec_high_county")
social_capital$`County Code`<- as.integer(social_capital$`County Code`)
social_capital$Population<- as.integer(social_capital$Population)
social_capital<- social_capital %>% 
  separate(`County Name`, c("State", "County"))
social_capital
```

Groups social networkings vs Income:
```{r}
ggplot(data = social_capital) + geom_point(aes(x=`High Income`, y = `Groups EC`))
```

Economic Connectedness vs Income:
```{r}
ggplot(data = social_capital) + geom_point(aes(x=`High Income`, y = `EconomicConnectedness`))
```



Tidying primary energy report:
```{r}
countries <- primary_energy %>% 
  filter(is.na(API)) %>% 
  rename(c("Country" = "...2"))
countries

production <- primary_energy %>% 
  filter(...2 == "Production (quad Btu)") %>% 
  rename(c("Production" = "...2")) %>% 
  select(2:42) %>% 
  mutate_if(is_character, as.double) %>% 
  round(3) %>% 
  mutate(Country = countries$Country) %>% 
  select(Country, 2:41) 
production
```

Primary Energy - COAL
```{r}
coal <- primary_energy %>%                               # COAL
  filter(...2 == "Coal (quad Btu)") %>%
  rename(c("Primary Energy" = "...2")) %>% 
  select(3:42) %>% 
  mutate_if(is_character, as.double) %>% 
  round(3)
coal <- coal
for(i in seq(1, nrow(coal), 2)){
  coal[i,]<- (coal[i,] + coal[i+1,])/2
}
coal <- coal %>% 
  filter(row_number() %% 2 != 0) %>%
  round(3) %>% 
  mutate(Country = countries$Country) %>%
  mutate(PrimaryEnergy = c("Coal (quad Btu)")) %>% 
  select(PrimaryEnergy, Country, 1:40)
coal
```

Primary Energy - NATURAL GAS
```{r}
natural_gas <- primary_energy %>% 
  filter(...2 == "Natural gas (quad Btu)") %>% 
  rename(c("Primary Energy" = "...2")) %>% 
  select(3:42) %>% 
  mutate_if(is_character, as.double) %>% 
  round(3)
natural_gas <- natural_gas
for(i in seq(1, nrow(natural_gas), 2)){
  natural_gas[i,]<- (natural_gas[i,] + natural_gas[i+1,])/2
}
natural_gas <- natural_gas %>% 
  filter(row_number() %% 2 != 0) %>%
  round(3) %>% 
  mutate(Country = countries$Country) %>%
  mutate(PrimaryEnergy = c("Natural Gas (quad Btu)")) %>% 
  select(PrimaryEnergy, Country, 1:40)
natural_gas
```

Primary Energy - CONSUMPTION
```{r}
consumption <- primary_energy %>% 
  filter(...2=="Consumption (quad Btu)") %>% 
  rename(c("Consumption" = "...2")) %>% 
  select(3:42) %>% 
  mutate_if(is_character, as.double) %>% 
  round(3) %>% 
  mutate(Country = countries$Country) %>% 
  select(Country, 1:40)
consumption
```

Primary Energy - PETROLEUM
```{r}
petroleum <- primary_energy %>% 
  filter(...2=="Petroleum and other liquids (quad Btu)") %>% 
  rename(c("Petroleum" = "...2")) %>% 
  select(3:42) %>% 
  mutate_if(is_character, as.double) %>% 
  round(3)
petroleum <- petroleum
for(i in seq(1, nrow(petroleum), 2)){
  petroleum[i,]<- (petroleum[i,] + petroleum[i+1,])/2
}
petroleum <- petroleum %>% 
  filter(row_number() %% 2 != 0) %>%
  round(3) %>% 
  mutate(Country = countries$Country) %>%
  mutate(PrimaryEnergy = c("Petroleum and Other Liquids")) %>% 
  select(PrimaryEnergy, Country, 1:40)
petroleum
```

Primary Energy - NUCLEAR
```{r}
nuclear_renewables <- primary_energy %>% 
  filter(...2 == "Nuclear, renewables, and other (quad Btu)") %>% 
  select(3:42) %>% 
  mutate_if(is_character, as.double) %>% 
  round(3)
for(i in seq(1, nrow(nuclear_renewables), 2)){
  nuclear_renewables[i,]<- (nuclear_renewables[i,] + nuclear_renewables[i+1,])/2
}
nuclear_renewables <- nuclear_renewables %>% 
  filter(row_number() %% 2 != 0) %>%
  round(3) %>% 
  mutate(Country = countries$Country) %>%
  mutate(PrimaryEnergy = c("Nuclear, renweables, and other")) %>% 
  select(PrimaryEnergy, Country, 1:40)
nuclear_renewables
```


```

